---
title: "lab3"
author: "Claire Madden"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
#set options that you want to apply to the whole markdown doc here! can override in a specific code chunk later on if you want to show specific things - code chunk header settings will override global settings here


library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
```

## Binary Logistic Regression: Penguins!

Read in the data:
```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

# if in doubt, the most complete way to join dfs is full join - will keep everything, and then you can wrangle to get the subset you want (other join options will exclude things based on whether the other dataset has a match or not which could cause lost data!)

penguins <- full_join(chinstrap, gentoo) %>% 
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% 
  filter(sex %in% c("male", "female")) #%in% operator is a way of doing a more elegant "for" statement and also great if you want to keep in a lot of stuff, another way to do it would be filter(sex=="male" | sex = "female") but this would take too long if there were a lot of matching options we were trying to keep in the df


```

Now, let's graph it!
```{r}

ggplot(data = penguins, aes(x = sex, y = body_mass_g))+
  geom_jitter(
    aes(color = sex, 
        pch = sex)
  )+
  facet_wrap(~species)


```


Code chinstrap penguins as (1) and gentoo penguins as (0):

```{r}
# could think about an ifelse statement
# case_when is built to do ifelse statements 
# in case_when, ~ means "then"
# if we had more than two species and wanted to assign all others the same thing, can use T to mean "if anything else is true" sort of like an else statement, but would have to be another numeric because other inputs are numeric

penguin_bin <- penguins %>% 
  mutate(
    sp_binary = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1, 
      species == "Gentoo penguin (Pygoscelis papua)" ~ 0
      )
    )


```


Now let's run our binary logistic regression:

```{r}

penguin_blr <- glm(sp_binary ~ sex + body_mass_g,
                   family = binomial,
                   data = penguin_bin)


summary(penguin_blr)
# coefficient estimates shows information for chinstrap because that was coded (1)
# as body mass increases, it becomes less and less probable that the penguin is a chinstrap
# still in terms of logodds which is difficult to interpret 
```


Let's make some predictions for penguins (classification, kinda):

1. What is the probability that a penguin is a chinstrap penguin, if it weighs 4500 grams and is male?

```{r}
# create a new dataframe that contains only the variables that we are interested in (asked in the question), variable name and entry names need to match exactly what the regression model is expecting

df_m4500 <- data.frame(sex = "male",
                   body_mass_g = 4500)

# next we make a prediction using the df we just created

# find the log odds of this penguin being chinstrap

m_4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link")

m_4500_logodds

#type = "link" reports the log odds, kinda hard to interpret

m_4500_prob <- predict(penguin_blr, newdata = df_m4500, type = "response")

m_4500_prob

#type = "response" reports the probability 
```

2. what is the probability that a penguin is chinstrap if it is recorded as female and 4000g?

```{r}

df_f4000 <- data.frame(sex = "female",
                       body_mass_g = 4000)



f_4000_prob <- predict(penguin_blr, newdata = df_f4000, type = "response")

f_4000_prob

```

Now let's make an entire data frame and use it to make then visualize logistic regression outcomes:


```{r}

# repeat sequence twice, once for male, once for female
# whatever length the first column is specified to, need to be sure other columns have the same number of rows 
penguins_mock <- data.frame(
  body_mass_g = rep(seq(3000,6000, length = 200), 2),
  sex = c(rep("male", 200), rep("female", 200))
)


```

Now let's use those to make predictions for all of them:

```{r}
full_predict <- predict(penguin_blr, newdata = penguins_mock, type = "response", se.fit = TRUE)

# get those predictions into a dataframe along with the mock data used to create them

final_df <- data.frame(
  penguins_mock,
  full_predict$fit,
  full_predict$se.fit
)

colnames(final_df) <- c("penguin_mass", "sex", "probability", "se")


```

Now, let's graph it!

```{r}
ggplot(data = final_df, aes(x = penguin_mass, y = probability))+
  geom_line(aes(color = sex))+
  geom_ribbon(aes(ymin = probability - se, 
                  ymax = probability + se, 
                  fill = sex),
              alpha = 0.3)



```


## Now some spatial data with sf

Read in the spatial data using read_sf
```{r}
cougars <- read_sf(dsn = here("cougar_connections"), layer = "ds1014") %>% 
  clean_names()

```



```{r}
# to demonstrate sticky geometry in spatial data, when you select area_ac you also keep the geometry 

large_corridors <- cougars %>% 
  filter(area_ac > 20000) %>% 
  select(area_ac)

plot(large_corridors)
```

```{r}

cougars_sub<-cougars %>% 
  select(elev_mean)


```


in ggplot:

```{r}
ggplot(data = cougars_sub)+
  geom_sf(aes(fill = elev_mean),
          color = NA)
```


```{r}
# tmap default is static map, override by mode ("view") to make interactive

tmap_mode("view")

tm_shape(cougars_sub)+
  tm_fill("elev_mean")

```






